# Linear morphometrics

```{r load packages 3, echo=TRUE, warning=FALSE}
# install required analysis packages
#devtools::install_github("tidyverse/tidyverse")
#devtools::install_github("mlcollyer/RRPP")
#devtools::install_github("kassambara/ggpubr")
#devtools::install_github("sinhrks/ggfortify")
#devtools::install_github("daattali/ggExtra")
# load libraries
library(here)
library(tidyverse)
library(RRPP)
library(ggpubr)
library(ggfortify)
library(cluster)
library(wesanderson)
library(ggExtra)
```

## Read data and define variables

```{r setwd ld define variables 3, echo=TRUE}
# read data
data <- read.csv("qdata.csv", header = TRUE, as.is=TRUE)

# define variables
maxl <- data$maxl # maximum length
maxw <- data$maxw # maximum width
maxth <- data$maxth # maximum thickness
maxstl <- data$maxstl # maximum stem length
maxstw <- data$maxstw # maximum stem width
raw.mat <- data$raw.mat # raw material
```

## Boxplots for _variable_ by _raw material_

```{r boxraw.mat, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap = "Boxplots for maximum length, width, thickness, stem length, and stem width for Perdiz arrow points."}
# boxplot of maximum length ~ raw.mat
raw.matmaxl <- ggplot(data, aes(x = raw.mat, y = maxl, color = raw.mat)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  theme(legend.position = "none") +
  labs(x = 'Raw Material', y = 'MaxL')

# boxplot of maximum width ~ raw.mat
raw.matmaxw <- ggplot(data, aes(x = raw.mat, y = maxw, color = raw.mat)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  theme(legend.position = "none") +
  labs(x = 'Raw Material', y = 'MaxW')

# boxplot of maximum thickness ~ raw.mat
raw.matmaxth <- ggplot(data, aes(x = raw.mat, y = maxth, color = raw.mat)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  theme(legend.position = "none") +
  labs(x = 'Raw Material', y = 'MaxTh')

# boxplot of stem length ~ raw.mat
raw.matmaxstl <- ggplot(data, aes(x = raw.mat, y = maxstl, color = raw.mat)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  theme(legend.position = "none") +
  labs(x = 'Raw Material', y = 'StemL')

# boxplot of stem width ~ raw.mat
raw.matmaxstw <- ggplot(data, aes(x = raw.mat, y = maxstw, color = raw.mat)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  theme(legend.position = "none") +
  labs(x = 'Raw Material', y = 'StemW')

# render figure
raw.matfigure <- ggarrange(raw.matmaxl, raw.matmaxw, raw.matmaxth, raw.matmaxstl, raw.matmaxstw,
                  labels = c("a","b","c","d","e"),
                  ncol = 2, nrow = 3)
              
## plot figure
raw.matfigure
```

## Principal Components Analysis

```{r pcaraw.mat, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap = "Principal components analysis by raw material."}
# attributes for plot
df<-data[c(4:8)]
pch.gps.gp <- c(15,16)[as.factor(raw.mat)]
col.gps.gp <- wes_palette("Moonrise2")[as.factor(raw.mat)]

## pca plot
pca <- autoplot(prcomp(df),
                data = data,
                asp = 1,
                shape = pch.gps.gp,
                colour = "raw.mat",
                variance_percentage = TRUE,
                loadings = TRUE, 
                loadings.colour = 'blue',
                loadings.label = TRUE,
                loadings.label.size = 3,
                frame = TRUE,
                frame.type = 't') +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2"))

ggMarginal(pca, groupColour = TRUE)
```

## Analyses of Variance (ANOVA) for _variable_ ~ _raw.mat_

### _Maximum length_ of Perdiz arrow points

```{r anovaraw.matmaxl, out.width = "100%", dpi = 300, echo=TRUE}
# anova = maximum length ~ raw.mat
raw.matml <- lm.rrpp(maxl ~ raw.mat, 
                  SS.type = "I", 
                  data = data, iter = 9999, 
                  print.progress = FALSE)
anova(raw.matml)
```

### _Maximum width_ of Perdiz arrow points

```{r anovaraw.matmaxw, out.width = "100%", dpi = 300, echo=TRUE}
# anova = maximum width ~ raw.mat
raw.matmw <- lm.rrpp(maxw ~ raw.mat, 
                  SS.type = "I", 
                  data = data, 
                  iter = 9999, 
                  print.progress = FALSE)
anova(raw.matmw)
```

### _Maximum thickness_ of Perdiz arrow points

```{r anovaraw.matmaxth, out.width = "100%", dpi = 300, echo=TRUE}
# anova = maximum thickness ~ raw.mat
raw.matmth <- lm.rrpp(maxth ~ raw.mat, 
                   SS.type = "I", 
                   data = data, 
                   iter = 9999, 
                   print.progress = FALSE)
anova(raw.matmth)
```

### _Maximum stem length_ of Perdiz arrow points

```{r anovaraw.matmaxstl, out.width = "100%", dpi = 300, echo=TRUE}
# anova = maximum stem length ~ raw.mat
raw.matmstl <- lm.rrpp(maxstl ~ raw.mat, 
                    SS.type = "I", 
                    data = data, 
                    iter = 9999, 
                    print.progress = FALSE)
anova(raw.matmstl)
```

### _Maximum stem width_ of Perdiz arrow points

```{r anovaraw.matmaxstw, out.width = "100%", dpi = 300, echo=TRUE}
# anova = maximum stem width ~ raw.mat
raw.matmstw <- lm.rrpp(maxstw ~ raw.mat, 
                    SS.type = "I", 
                    data = data, 
                    iter = 9999, 
                    print.progress = FALSE)
anova(raw.matmstw)
```
