# Symmetry and asymmetry

In studies of projectile morphology, an analysis of bilateral/object symmetry may yield useful insights related to _design intent_, provided we assume that the goal of the knapper was to achieve a fully-symmetrical form. Within that framework, deviation from bilateral symmetry (fluctuating asymmetry) might serve as a useful measure of _knapper skill_ and/or _differential production_. Both measures have utility in comparisons of contexts and sites, to include wide-ranging spatial and temporal applications. Initially developed for a [symmetry analysis of Gahagan bifaces](https://github.com/aksel-blaise/gahagan.asymmetry/blob/master/asymmetry.md), the same approach can be employed for an analysis of asymmetry in Perdiz arrow points.

## Load packages + data

```{r load.data.asymm, echo=TRUE, out.width = "100%", dpi = 300, warning=FALSE}
# load analysis packages
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

# read shape data and define number of sLMs
shapes <- readShapes("shapes")
shapesGM <- readland.shapes(shapes, nCurvePts = c(10,3,5,5,3,10))

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)
land.pairs <- read.csv("asymmetry1.csv",
                     header=TRUE)
ind <- qdata$spec.x
```

## Generalized Procrustes Analysis

```{r assym.gpa, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# gpa for bilateral (object) asymmetry
res.bilat <- bilat.symmetry(shapesGM,
                          ind = ind, 
                          object.sym = TRUE,
                          land.pairs = land.pairs,
                          RRPP = TRUE,
                          iter = 9999,
                          print.progress = FALSE)
summary(res.bilat)
```

### Symmetric shape

```{r symm.shape, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Symmetric shape component."}
# symmetric shape
symm.shape <- res.bilat$symm.shape[,1:2,]

## plot all specimens
plotAllSpecimens(symm.shape)
```

### Asymmetric shape

```{r asymm.shape, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Asymmetric shape component."}
# asymmetric shape
asymm.shape <- res.bilat$FA.component[,1:2,]

## plot all specimens
plotAllSpecimens(asymm.shape)
```

## Principal Components Analysis (symmetric)

```{r pca.symm, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# geomorph data frame
gdf <- geomorph.data.frame(shape = res.bilat$symm.shape, 
                           ashape = res.bilat$FA.component,
                           raw.mat = qdata$raw.mat)

# principal components analysis (symmetric)
pca.symm <- gm.prcomp(res.bilat$symm.shape)
summary(pca.symm)

# set plot parameters
pch.gps <- c(15:17)[as.factor(raw.mat)]
col.gps <- pal[as.factor(raw.mat)]
col.hull <- c("#C27D38","#798E87","#CCC591")

## pca.symm plot
pc.plot <- plot(pca.symm,
                asp = 1,
                pch = pch.gps,
                col = col.gps)
shapeHulls(pc.plot,
           groups = raw.mat,
           group.cols = col.hull)
```

## Principal Components Analysis (asymmetric)

```{r pca.asymm, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# principal components analysis (asymmetric)
pca.asymm <- gm.prcomp(res.bilat$FA.component)
summary(pca.asymm)

## pca.asymm plot
pc.plot <- plot(pca.asymm,
                asp = 1,
                pch = pch.gps,
                col = col.gps)
shapeHulls(pc.plot,
           groups = raw.mat,
           group.cols = col.hull)
```

## Procrustes ANOVA

```{r def.mod.1.asymm, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# symmetric shape as a function of raw material
fit.shape.raw <- procD.lm(shape ~ raw.mat, 
                            data = gdf, 
                            print.progress = FALSE, 
                            iter = 9999)
# anova
anova(fit.shape.raw)

# asymmetric shape as a function of raw material
fit.shape.raw.asymm <- procD.lm(ashape ~ raw.mat, 
                            data = gdf, 
                            print.progress = FALSE, 
                            iter = 9999)

# anova
anova(fit.shape.raw.asymm)

# end of code
```
